A VLT/FORS2 Narrowband Imaging Search for \ion{Mg}{2} Emission Around $\lowercase{z}\sim0.7$ Galaxies

### 2018-06-06 Manuscript/vaught_2017.tex
updated paper May29th
updated June6th

\begin{document}
\title{A VLT/FORS2 Narrowband Imaging Search for \ion{Mg}{2} Emission Around $\lowercase{z}\sim0.7$ Galaxies }
\author{Ryan Rickards Vaught}
\affiliation{San Diego State University 5500 Campanile Dr, San Diego, CA 92182}
\affiliation{Department of Physics, University of California, San Diego, CA}

 \author{Kate H. R. Rubin }
 \affiliation{San Diego State University 5500 Campanile Dr, San Diego, CA 92182}

@@

\author{Joseph Hennawi,}
\affiliation{Department of Physics, Broida Hall, University of California, Santa Barbara, CA 93106-9530}

\correspondingauthor{Ryan Rickards Vaught}
\email{rjvaught@gmail.com}

@@

Galactic-scale outflows are thought to be the primary mechanism in the removal of cool gas from star-forming galaxies. Presently, the mass and energy of these flows remain poorly constrained. One way to better constrain these parameters is to measure the spatial extent of the outflow; however, measuring this quantity via spectral methods has been traditionally very difficult due to the faintness of emission lines tracing outflowing material. We perform a VLT/FORS2 narrowband imaging of 5 star-forming galaxies at redshift $z=0.67-0.69$ in the GOODS-S field as part of an effort to spatially resolve and constrain the radial extent of large-scale outflows traced by \ion{Mg}{2} emission. These observations probe \ion{Mg}{2} to unmatched surface brightness limits of 5.74 $\times$ $10^{-19}$ ergs sec $^{-1}$ cm$^{-2}$ arcsec$^2$ (5$\sigma$). We do not detect any extended \ion{Mg}{2} emission in any of our galaxies, suggesting that the wind material does not extend beyond the galaxies' stellar component. Our observations also allow us to create the first ever spatially-resolved map of  \ion{Mg}{2} absorption, revealing approximately constant absorption strengths across the galaxy disks.

@@

\keywords{Galaxy Evolution, Galactic Winds}

@@

To find the amount of missing emission, we first need to calculate how much continuum has been absorbed by \ion{Mg}{2}. Using our Keck/LRIS spectra, we find the average value of the continuum near the \ion{Mg}{2} doublet and multiply this value by the EW of the doublet. Although we know how much emission is missing, we do not know how this emission is distributed outside the galaxy. As a conservative estimate of the SB, we distribute the missing emission uniformly inside multiple and increasingly sized annuli. These annuli are fixed to have an inner radius equal to the galaxy's isophotal radius and an increasing outer radius.  Additionally, since our SB limits are dependent on the size of the aperture used, we calculate the SB limits inside each of the aforementioned annuli. Figure \ref{fig.emission} shows how the SB of emission varies with the spatial extent of the annulus, as well as how the SB compares with our detection limits..

Excluding the galaxy J033230.03, the SB of this missing emission lie above our detection limits. This result shows that these galaxies do not host an isotropic dust-free wind.

@@

\subsection{Morphology of MgII Absorption}

@@

\section{CONCLUSION AND FUTURE WORK}
We presented the results from a narrowband imaging search for \ion{Mg}{2} emission in a sample of star-forming galaxies at a redshift of $z \sim 0.70$.  Although we were unable to detect any measurable amount of \ion{Mg}{2} emission, we are able to report a conservative $5\sigma$ detection limit of  SB(\ion{Mg}{2}) = 5.74 $\times 10^{-19}$ ergs sec $^{-1}$ cm$^{-2}$ arcsec$^2$. We detected a significant amount of \ion{Mg}{2} absorption in a total of 4 galaxies out of our 5 galaxy sample. Furthermore, our imaging allows us to generate spatially-resolved maps of \ion{Mg}{2} absorption in a distant galaxy sample for the first time. This absorption covered the center of the galaxies out to the SB$_1$ continuum + \ion{Mg}{2} isophotal radius, at approximately $\sim 22$ kpc, suggesting that the absorbing gas fully covers the stellar disks out to this distance. The measured EWs measured in these maps are in broad agreement with measured EWs using Keck/LRIS integrated slit spectroscopy. Additionally, our radial projections of the mean EW for our sample galaxies suggest that the EWs due to \ion{Mg}{2} are approximately constant out towards the edges of the galaxies. Finally, we compared our detection limits with estimated SB \ion{Mg}{2} emission, based on the radiative transfer models of \cite{Prochaska_2011}. For an isotropic dust-free wind, \ion{Mg}{2} is predicted to be generate alongside ubiquitous \ion{Mg}{2} absorption. Our detection limits are suitable to detect the estimated SB \ion{Mg}{2} emission and since we do not detect the excess emission, the  isotropic morphology of a wind is ruled out in our galaxy sample.

In the future, we hope to observe these galaxies using an Integral Field spectrograph (IFU). IFUs take the signal from an individual pixel and feeds it into a spectrograph. The spectrograph then generates a spectrum for that pixel, allowing us to view the flux at specific transitions. IFUs cover large 2D fields of view, contrary to single point 1D spectroscopy or the slice of 2D longslit spectroscopy, makes them exceptionally useful in resolving the spectrum of a target for multiple sightlines. Thus, IFUs are useful in detecting the gas of extended objects, such as galaxies. Because IFUs would allow us to observe the spectrum of a galaxy's extended region, we will be able to compare our EW maps, as well as EW radial projections, with higher resolution observations.

Based on the radial extent and strength of our measured \ion{Mg}{2} absorption, the \ion{Mg}{2} is optically thick. However, we need to compare our measured absorption strength and radial extent with radiative transfer models. In the radiative transfer models of \cite{Prochaska_2011}, scattered emission is predicted for an isotropic, optically thick,and dust free wind with density that decreases as an inverse square law, however galaxies are not spherically symmetric and sources driving the wind may not be distributed spherically. Allowing for a bi-conical anisotropic wind, \cite{Prochaska_2011} state that the scattered \ion{Mg}{2} emission will become suppressed compared to the strengths an isotropic wind. Future observations of our galaxies in IFUs could perhaps shed light on this alternate scenario.

@@

\begin{table*}[]
\centering
\caption{Properties of \ion{Mg}{2}\ Absorption\label{tab:abs_props}}  
\begin{tabular}{llllll} \hline \hline
Object & Max(SB$_{abs}$)\footnote{SB Values are in units of $10^{-18}$ ergs sec $^{-1}$ cm$^{-2}$ arcsec$^2$} & $R_{\text{SB}_1}(kpc)$ &\ \ \ EW($\AA$)\footnote{ Measured from EW images} & EW($\AA$)\footnote{Measured from Keck/LRIS spectra}  \\  \hline
J033225.26-274524.0 &  $-5.39 \pm 1.28 $ & 8 &     $\ \ \ 3.511 \pm 0.795$ & $7.539 \pm 0.354 $\\
J033232.36-274725.0 &  $-5.64 \pm 0.121 $ & 15 & $\ \ \ 7.694 \pm 1.016$ & $5.835 \pm 0.493$\\
J033230.03-274347.3 &  $-4.41 \pm 0.622 $ & 21 & $\ \ \ 5.388 \pm 0.712$ & $12.79 \pm 1.710$\\
J033229.64-274242.5 &  $-18.2 \pm 0.127 $ & 10 & $\ \ \ 7.546 \pm 0.660$ & $13.24 \pm 0.263$\\
J033230.57-274518.2 &  $-1.25 \pm 1.15   $ & 11& -$0.6532 \pm 0.589$ & $6.106 \pm 0.370$\\ \hline
\end{tabular}
\end{table*}

\begin{figure*}
\centering
\gridline{\fig{../Figures/geo_J26.pdf}{0.5\textwidth}{(J033225.26)}
          \fig{../Figures/geo_J36.pdf}{0.5\textwidth}{(J033231.36)}}
\gridline{\fig{../Figures/geo_J03.pdf}{0.5\textwidth}{(J033230.03)}
          \fig{../Figures/geo_J64.pdf}{0.5\textwidth}{(J033230.64)}}
\caption{ Estimated SB profile of \ion{Mg}{2} emission. The square points are SB of missing \ion{Mg}{2} emission uniformly distributed inside an annulus. The solid line shows the value of the SB detection limits. The dashed vertical and horizontal lines represent outer radius of }
\label{fig.emission}
\end{figure*}

>>>>

\begin{document}
\title{A VLT/FORS2 Narrowband Imaging Search for \ion{Mg}{2} Emission Around $\lowercase{z}\sim0.7$ Galaxies }
\author{Ryan Rickards Vaught}
\affiliation{Department of Astronomy, San Diego State University, San Diego, CA 92182}
\affiliation{Center for Astrophysics and Space Sciences, Department of Physics, University of California, San Diego, CA}

 \author{Kate H. R. Rubin }
 \affiliation{Department of Astronomy, San Diego State University, San Diego, CA 92182}

@@

\author{Joseph Hennawi}
\affiliation{Department of Physics, Broida Hall, University of California, Santa Barbara, CA 93106-9530}

\correspondingauthor{Ryan Rickards Vaught}
\email{rjrickar@ucsd.edu}

@@

The mass and energy of galactic winds remain poorly constrained via traditional absorption line studies. One way to better constrain these parameters is to measure the spatial extent of the outflow in emission. We perform a VLT/FORS2 narrowband imaging of 5 star-forming galaxies at redshift $z=0.67-0.69$ in the GOODS-S field as part of an effort to spatially resolve and constrain the radial extent of large-scale outflows traced by \ion{Mg}{2} emission. These observations probe extended \ion{Mg}{2} emission down to unmatched surface brightness limits of 5.74 $\times$ $10^{-19}$ ergs sec $^{-1}$ cm$^{-2}$ arcsec$^2$ (5$\sigma$).  We do not detect any extended \ion{Mg}{2} emission in any of our galaxies, thus placing strong upper limits on the strength of \ion{Mg}{2} emission in the sample galaxies. Our observations allow us to create the first ever spatially-resolved map of \ion{Mg}{2} absorption, revealing approximately constant absorption strengths across the galaxy disks. Our detection limits in concert with a previous studies Keck/LRIS spectra allow us to model \ion{Mg}{2} emission for isotropic and anisotropic dust-free winds. We find that anisotropic winds reduce the strength of estimated missing \ion{Mg}{2} emission below our 5$\sigma$ limit.

@@

\keywords{galaxies: evolution, Galactic Winds}

@@

To find the amount of missing emission, we first calculate how much continuum has been absorbed by \ion{Mg}{2}. Using our Keck/LRIS spectra, we find the average value of the continuum near the \ion{Mg}{2} doublet and multiply this value by the EW of the doublet. Although we know how much emission is missing, we do not know how this emission is distributed outside the galaxy. As a conservative estimate of the SB, we distribute the missing emission uniformly inside multiple and increasingly sized annuli. These annuli are fixed to have an inner radius equal to the galaxy's isophotal radius and an increasing outer radius.  Additionally, since our SB limits are dependent on the size of the aperture used, we calculate the SB detection limits of our images inside each of the aforementioned annuli. Figure \ref{fig.emission} shows how the SB of emission varies with the spatial extent of the annulus, as well as how the SB compares with our detection limits. Excluding the galaxy J033230.03, the SB of this missing emission lie above our detection limits. This result shows that these galaxies do not host an isotropic dust-free wind.

There are many phenomena that we can employ to reduce the SB of the missing \ion{Mg}{2} emission to better match our observations. Introducing dust into the wind can reduce the observed emission as it can change the observed line profile in a critical way as re-emitted photons have to travel farther distances to escape the wind and therefore the profile will suffer greater extinction. Another factor that can reduce the amount of emission is anisotropy of the wind. Isotropic winds are idealized, but galaxies are not spherically symmetric and the shape of the galaxies interstellar medium may preferentially suppress outflows at varying strengths for different galactic longitudes. We now assume that the emission in our galaxies is reduced by purely by the effects anisotropy, as the constraints on dust attenuation or unknown. For the anisotropic winds modeled in \cite{Prochaska_2011}, the factor that the emission is reduced is given by the scaling $\Omega/4\pi$, where $\Omega$ is the angular extent of the wind which is constrained to point towards the observer $\Omega > 2\pi$. We show the anisotropic wind SB profiles in Figure  \ref{fig.emission}.

After reducing the SB of the missing \ion{Mg}{2} emission by the minimum opening angle $\Omega=2\pi$, we are able to profiles that fall below our SB detection limits for galaxy's J033225.26 and J033231.36. However, the SB profile of J033229.64 remains above our SB detection limits, suggesting additional phenomena are needed to reduce the strength of missing emission.

@@

\begin{table*}[]
\centering
\caption{Properties of \ion{Mg}{2}\ Absorption\label{tab:abs_props}}  
\begin{tabular}{llllll} \hline \hline
Object & Max(SB$_{abs}$)\footnote{SB Values are in units of $10^{-18}$ ergs sec $^{-1}$ cm$^{-2}$ arcsec$^2$} & $R_{\text{SB}_1}(kpc)$ &\ \ \ EW($\AA$)\footnote{ Measured from EW images} & EW($\AA$)\footnote{Measured from Keck/LRIS spectra}  \\  \hline
J033225.26-274524.0 &  $-5.39 \pm 1.28 $ & 8 &     $\ \ \ 3.511 \pm 0.795$ & $7.539 \pm 0.354 $\\
J033232.36-274725.0 &  $-5.64 \pm 0.121 $ & 15 & $\ \ \ 7.694 \pm 1.016$ & $5.835 \pm 0.493$\\
J033230.03-274347.3 &  $-4.41 \pm 0.622 $ & 21 & $\ \ \ 5.388 \pm 0.712$ & $12.79 \pm 1.710$\\
J033229.64-274242.5 &  $-18.2 \pm 0.127 $ & 10 & $\ \ \ 7.546 \pm 0.660$ & $13.24 \pm 0.263$\\
J033230.57-274518.2 &  $-1.25 \pm 1.15   $ & 11& -$0.6532 \pm 0.589$ & $6.106 \pm 0.370$\\ \hline
\end{tabular}
\end{table*}
\subsection{Morphology of MgII Absorption}

@@

\section{CONCLUSION}
We presented the results from a narrowband imaging search for \ion{Mg}{2} emission in a sample of star-forming galaxies at a redshift of $z \sim 0.70$.  Although we were unable to detect any measurable amount of \ion{Mg}{2} emission, we are able to report a conservative $5\sigma$ detection limit of  SB(\ion{Mg}{2}) = 5.74 $\times 10^{-19}$ ergs sec $^{-1}$ cm$^{-2}$ arcsec$^2$. We detected a significant amount of \ion{Mg}{2} absorption in a total of 4 galaxies out of our 5 galaxy sample. Furthermore, our imaging allows us to generate spatially-resolved maps of \ion{Mg}{2} absorption in a distant galaxy sample for the first time. This absorption covered the center of the galaxies out to the SB$_1$ continuum + \ion{Mg}{2} isophotal radius, at approximately $\sim 22$ kpc, suggesting that the absorbing gas fully covers the stellar disks out to this distance. The measured EWs measured in these maps are in broad agreement with measured EWs using Keck/LRIS integrated slit spectroscopy. Additionally, our radial projections of the mean EW for our sample galaxies suggest that the EWs due to \ion{Mg}{2} are approximately constant out towards the edges of the galaxies.

Finally, we compared our detection limits with estimated SB \ion{Mg}{2} emission, based on the radiative transfer models of \cite{Prochaska_2011}. We are able to rule out that the winds in our sample are isotropic, as our detection limits are suitable to detect the predicted emission. By assuming anisotropic winds, we were able to reduce the strength of the predicted emission to lie below our detection limit, albeit one galaxy that needs additional phenomena like dust to achieve non-detectability in our images.

Based on the radial extent and strength of our measured \ion{Mg}{2} absorption, the \ion{Mg}{2} is optically thick. Deeper images in \ion{Mg}{2} are required to observe any extended emission beyond the depth of our images. Although we are able to rule out that the winds in our sample are isotropic and dust-free, questions linger on weather the wind is anisotropic as well as dust-free. A new generation of instruments such as the Keck Cosmic Web Imager (KCWI) or the Multi Unit Spectroscopic Explorer (MUSE) should be used to answer the prevailing questions

%In the future, we hope to observe these galaxies using an Integral Field spectrograph (IFU). IFUs take the signal from an individual pixel and feeds it into a spectrograph. The spectrograph then generates a spectrum for that pixel, allowing us to view the flux at specific transitions. IFUs cover large 2D fields of view, contrary to single point 1D spectroscopy or the slice of 2D longslit spectroscopy, makes them exceptionally useful in resolving the spectrum of a target for multiple sightlines. Thus, IFUs are useful in detecting the gas of extended objects, such as galaxies. Because IFUs would allow us to observe the spectrum of a galaxy's extended region, we will be able to compare our EW maps, as well as EW radial projections, with higher resolution observations.

%Based on the radial extent and strength of our measured \ion{Mg}{2} absorption, the \ion{Mg}{2} is optically thick. However, we need to compare our measured absorption strength and radial extent with radiative transfer models. In the radiative transfer models of \cite{Prochaska_2011}, scattered emission is predicted for an isotropic, optically thick,and dust free wind with density that decreases as an inverse square law, however galaxies are not spherically symmetric and sources driving the wind may not be distributed spherically. Allowing for a bi-conical anisotropic wind, \cite{Prochaska_2011} state that the scattered \ion{Mg}{2} emission will become suppressed compared to the strengths an isotropic wind. Future observations of our galaxies in IFUs could perhaps shed light on this alternate scenario.

@@

\begin{figure*}[h]
\centering
\gridline{\fig{../Figures/geo_J26.pdf}{0.5\textwidth}{(J033225.26)}
          \fig{../Figures/geo_J36.pdf}{0.5\textwidth}{(J033231.36)}}
\gridline{\fig{../Figures/geo_J03.pdf}{0.5\textwidth}{(J033230.03)}
          \fig{../Figures/geo_J64.pdf}{0.5\textwidth}{(J033230.64)}}
\caption{Estimated SB profile of \ion{Mg}{2} emission for Isotropic and Anisotropic dust-free winds. The black square points are SB of missing \ion{Mg}{2} emission uniformly distributed inside an annulus for an angle $\Omega=4\pi$, or an isotropic wind. The blue diamond points are SB of missing \ion{Mg}{2} emission uniformly distributed inside an annulus for an angle $\Omega=2\pi$, or an anisotropic wind. The solid line shows the value of the SB detection limits. The dashed vertical and horizontal lines represent outer radius of the extended annulus use to measure the initial detection limits and the value of SB detection limit in that annulus.}
\label{fig.emission}
\end{figure*}
